---
title: "Week3"
output: pdf_document
date: "2023-09-12"
---
Question 5.1
Using crime data from the file uscrime.txt (http://www.statsci.org/data/general/uscrime.txt,
description at http://www.statsci.org/data/general/uscrime.html), test to see whether there are any
outliers in the last column (number of crimes per 100,000 people). Use the grubbs.test function in
the outliers package in R.

```{r}
#clean the R environment
rm(list = ls())
set.seed(10)

#load data into dataframe
crimedf <- read.table("uscrime.txt", stringsAsFactors = FALSE, header = TRUE)
head(crimedf)

```

```{r}
#load library
#install.packages("outliers")
library(outliers)
library(ggplot2)
library(tidyverse)
```

H0:There is no outlier in crime data.
```{r}
#using type = 10 to test for 1 outlier. test the maximum value
grubbs.test(crimedf$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
#test the minimum value
grubbs.test(crimedf$Crime, type = 10, opposite = TRUE, two.sided = FALSE)

#using type =11 to test for 2 outliers on opposite tails.
grubbs.test(crimedf$Crime, type = 11, opposite = FALSE, two.sided = FALSE)

```
As the result shown, the p-value = 0.07887 > alpha = 0.05. In this case, we don't have strong evidence to reject the null hypothesis. Therefore, we cannot find outlier in uscrime data.

Let's take a look at the box graph of the crime data. 
```{r boxgraph, echo=FALSE, fig.width=3}
#plot box graph to visualize finding
boxplot(crimedf$Crime,
        main = "Outlier for Crime in U.S. Crime",
        xlab = "Population Density",
        ylab = "Crime Committed",
        horizontal = FALSE,
        notch = FALSE)

```

Even the graph shown there are two outliers in the data set. Based on the p-value we found previously is 0.07887. The outliers maybe occur due the random noise. In the big data, outlier may appears more often than case study. But in the uscrime data, I would keep the outliers. The outlier may depends on the population in the states. Where the city or states has higher poplulation density, the higher chance to commit a crime. That can be the reason for the outliers.




#Question 6.1
Describe a situation or problem from your job, everyday life, current events, etc., for which a Change Detection model would be appropriate. Applying the CUSUM technique, how would you choose the critical value and the threshold?

Answer
We can refer the real life situation into subway maintenance, similar situation we can think of citi bike in New York City. This is a bike sharing project for city worker commute during rush hours. Using change detection model for monitor the repair for bike would be appropriate. For instance, New York City has about 2 millions bike trips everyday. We can use the change detection model to monitor the bike usage. We can calculate the approximate use time of each bike, trace the bike usage and detect the best time we should do repairs and maintenance for the bikes. 


#Question 6.2
1. Using July through October daily-high-temperature data for Atlanta for 1996 through 2015, use a CUSUM approach to identify when unofficial summer ends (i.e., when the weather starts cooling off) each year. You can get the data that you need from the file temps.txt or online, for example at http://www.iweathernet.com/atlanta-weather-records or https://www.wunderground.com/history/airport/KFTY/2015/7/1/CustomHistory.html. You can use R if you’d like, but it’s straightforward enough that an Excel spreadsheet can easily do the job too.

```{r}
#install.packages("lubridate")
library(repr)
library(reshape)

#load temp data
temp <- read.table("temps.txt", stringsAsFactors = FALSE, header = TRUE)
head(temp)

# average the temperature for each day across the years
avg_date_temp <- rowMeans(temp[c(2:length(temp))], na.rm=T)
# compute the mean of the average temperature by the same date from 1996-2015.
mu <- mean(avg_date_temp)
#find the standard deviation 
std_temp <- sd(avg_date_temp)
# set C
C <- std_temp

# set threshold T. 
T <- 4*std_temp

# create an empty column to store the data 
temp[,"St"]<-NA

# starts loop from 0.

temp[1,"St"]<-0
for(i in 1:nrow(temp)){
  
  temp[i,"St"]<-max(0,(temp[i-1,"St"]+mu-avg_date_temp[i]-C))
  
  
}

temp$St

cat("The day a change in trend is detected is:",temp[which(temp$St>T),"DAY"][1])

#change data type for "day" as month/date
temp[,"Date"]<-as.Date(temp[,"DAY"],"%d-%B")
temp[,"Date"]<-format(temp[,"Date"],format="%m/%d")


```


```{r control_graph, echo=FALSE}
#plot the control graph for visualization.
options(repr.plot.width=10, repr.plot.height=5)
ggplot(data = temp, aes(x = Date, y =`St`,group=2)) + 
  geom_line()+
  geom_hline(yintercept=T,color="red")+
  
  scale_x_discrete(breaks = unique(temp$Date)[seq(1,123,10)])+
  xlab("Days") +
  ylab("CUSUM") +
  ggtitle("CUSUM Chart for July-October Daily-high Temperature for Atlanta 1996-2015")+
  theme(plot.title = element_text(hjust = 0.5))

#As we can observed from the graph. On 10/18 the temperature starts to cool down. 

```
Per the control graph shown above, we can see the fall starts around 10/18 in Atlanta. The result is appropriate because geographically speaking. Atlanta located in the south part of the United States. Which means the fall starts a little later by the end of September until mid-October is applicable. 

2. Use a CUSUM approach to make a judgment of whether Atlanta’s summer climate has gotten warmer in that time (and if so, when).

In the approach, I used the cusum of each year to perform the model.

```{r}
#creating a data frame to store the cusum values for each year instead of same day on every year.
cusum_temp<-data.frame(matrix(nrow=nrow(temp),ncol=length(1996:2015)+1))

#assigning columns names to cusum data frame
colnames(cusum_temp)<-colnames(temp[,1:21])

#converting Day into Date that the loop can process easier.
cusum_temp$DAY<-temp$Date

# starts from zero
#calculating cusum values for each year
for(y in 2:ncol(cusum_temp)){
  cusum_temp[1,y]<-0 #initial St value for each column,set to zero
  mu<-mean(temp[,y]) #mean of each sample space(each year's observations)
  std<-sd(temp[,y]) #sd of each sample,also used as allowable slack
  threshold<-5*std #using 5 sd as threshold value,different T for each year 
  change<-NULL # to store dates with St over threshold,first value:first day change detected
  
  for(i in 2:nrow(cusum_temp)){
    cusum_temp[i,y]<-max(0,cusum_temp[i-1,y]+(mu-temp[i,y]-std))
    if (cusum_temp[i,y]>=threshold){
      change<-append(change,cusum_temp[i,y])}}
  cat("In ",colnames(cusum_temp[y])," first day of Fall started on",
      cusum_temp[which(cusum_temp[,y]==change[1]),"DAY"],"\n")
}
```

As we can see from the statements shown above, from 1996 to 2015 fall generally starts in October. For the giving data, we do not have strong evidence to proof that summer climate has gotten warmer. Perhaps, the data is not large enough to show the difference. In addition, even if some year may have longer summer time, that maybe just an "outlier" through out the time series. However, we cannot ignore the outlier because the exist of outlier is meaningful to help us study the climate.
