---
title: "Hw 2.2"
author: "Yawen Liang"
date: "2023-08-29"
output:
pdf_document: default
html_document:
df_print: paged
word_document: default
---

```
Question 2.1

Describe a situation or problem from your job, everyday life, current events, etc., for which a classification model would be appropriate. List some (up to 5) predictors that you might use.

I worked in an accounting company, and I review client’s business balance sheet as daily task. We would use some vectors such as revenue, cost of goods, business liabilities, owner's equity and assets. We can simply using revenue and cost of goods to calculate the profit that the company earns. Furthermore, besides analyze our client's business. We can also use some vectors such as employee retension rate, new hire rate, average salary with the same title in the company to predict the company transition rate. Also for the managers prespective, they can analyze based on the transition rate and retension rate to create better management strategy. 

```

---
#2.2.1

```{r}
library(kernlab)

#import data from txt.
data <- read.table("credit_card_data.txt", header = FALSE)

#define ksvm function. Call the linear equation to find the best C-value
model <- ksvm(as.matrix(data[,1:10]),as.factor(data[,11]),type='C-svc',kernel='vanilladot',C=10000,scaled=TRUE)

# calculate a1…am
a <- colSums(model@xmatrix[[1]] * model@coef[[1]])
a
# calculate a0
a0 <- -model@b
a0

# see what the model predicts
pred <- predict(model,data[,1:10])
pred

# see what fraction of the model’s predictions match the actual classification
sum(pred == data[,11]) / nrow(data)
```

#Test result: 1. C=100, accuracy is 0.8639144. Then changed the C-value to C=50,000. Trying to see if I changed a larger c-value will the accuracy rate be different from the first test. When C=50,000, the accuracy is 0.8623853.In my third test. I tested C=100,000. The results came out the accuracy rate dropped from 0.8623853 to 0.6253823.
#In conclusion, I decided to keep the cost between 100 - 10,000 since the results shown the cost within this range has the highest accuracy.

```{r}
#for loop for different c values
cvalues <- c(0.00001,0.0001,0.01,1,10,100,10000,1000000)
#setup linear equation for kernel we want to use in this question
k_kernel = 'vanilladot'
#create a dataframe to store the data from the for loop
results <- data.frame("C-Values" = integer(),"accuracy" = numeric())

for(cvalue in cvalues)
{
  #create a ksvm model to run for the test.
  model <- ksvm(as.matrix(data[,1:10]),as.factor(data[,11]),type='C-svc',kernel=k_kernel,C=cvalue,scaled=TRUE)
  
  #use ksvm to run formula and loop for different c value.
  pred <- predict(model,data[,1:10])
  
  #result in fraction of the model’s predictions match the actual classification
  accuracy <- sum(pred == data[,11]) / nrow(data)
  
  #create an empty list to store output values
  results[nrow(results) + 1,] <- c(cvalue,accuracy)
}
  # we find out the highest accuracy rate is 0.8639. Where the c value in between 0.01 - 100.
  #Therefore, the best c vaule of this model we can pick from 0.01-100 to get the highest accuracy rate.
  
View(results)
```

---
#2.2.2 (Optional)
```{r}
#try different methods other than vanilladot
diffkernels <- c("rbfdot","splinedot","laplacedot","polydot","vanilladot")

#for loop for different c values
cvalues <- c(0.00001,0.0001,0.01,1,10,100,10000,1000000)

#create a dataframe to store the data from the for loop
results <- data.frame("C-Values" = integer(),"accuracy" = numeric(),"kernel_name" = character(),stringsAsFactors = FALSE)

#n starts from 0.
n <- 0

for (ikernel in diffkernels) {

  n <- n+1  
  for(cvalue in cvalues)
  {
#setup linear equation for kernel we want to use in this question,loop for different kernels and c valuse.
    model <- ksvm(as.matrix(data[,1:10]),as.factor(data[,11]),type='C-svc',kernel=ikernel,C=cvalue,scaled=TRUE)
    
#use ksvm to run formular in different c value.
    pred <- predict(model,data[,1:10])
    
#result in fraction of the model’s predictions match the actual classification
    accuracy <- sum(pred == data[,11]) / nrow(data)
    
#need function to contain value
    results[nrow(results) + 1,] <- c(cvalue,accuracy,ikernel)
  }
}

#We find out the highest accuracy rate is 0.8639. Where the c value in between 0.01 - 100.The best model for this test is rbfdot. 
View(results)
```

---

---
#2.2.3
#install.package("kknn")
```{r}
library(kknn)

#import data
knndata <- read.table("credit_card_data.txt", header = FALSE,stringsAsFactors = FALSE)

head(knndata)

#create function to run for the train and test value, get the accuracy rate 
accuracy_check = function(X){
predicted <- rep(0,(nrow(data))) 
  
  for (i in 1:nrow(data))
  {
    
    #define model for calculation
    knn_credit=kknn(V11~.,knndata[-i,],knndata[i,],k=X, scale = TRUE) 
    
    predicted[i] <- as.integer(fitted(knn_credit)+0.5) # round off to 0 or 1
  }
  
  # calculate % of correct predictions
  accuracy<- sum(predicted == knndata[,11]) / nrow(data)
  return(accuracy)
}

# generate number from 1-20 of k values to test accuracy rate
kknn_accuracy<- rep(0,20)

for (i in 1:20)
{
  #looping k value
  kknn_accuracy[i] <- accuracy_check(i)
}

```

#As we can see from the graph, where K values in between 10-15 we can get the most accurate rate.

```{r pressure, echo=FALSE}
#Elbow Plot:  accuracy vs. K-values
par(bg="white")
plot(kknn_accuracy,ylab="accuracy level",xlab="K values" ,type='b',col='blue', main='Fig: KKN accuracy')
```

---

